---
title: "finalgroupproject"
author: "Tee Zhi Zhang"
date: "2023-11-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Final Group Project

### Group: Marcus Wan, Tan Xue Feng, Tee Zhi Zhang

## Introduction of Data

-   Price of HDB Resale Flat Sales: <https://beta.data.gov.sg/collections/189/datasets/d_8b84c4ee58e3cfc0ece0d773c8ca6abc/view>

-   Price of HDB BTO Launches:

    <https://beta.data.gov.sg/datasets/d_2d493bdcc1d9a44828b6e71cb095b88d/view> 

```{r}
library(tidyverse)
library(dplyr)
library(lubridate)
```

## Data Cleaning and Preprocessing for Resale Flat Prices

-   Remove street name, lease commence date, flat model (marcus)

Our data set is split into 4 different csv files which contain data for different time periods. Let us read the data examine the columns and data types.

```{r}
resale_data_19901999 <- read.csv('resale_19901999.csv', header = TRUE, sep = ',', encoding = 'UTF-8')
resale_data_20002012 <- read.csv('resale_2000Feb2012.csv', header = TRUE, sep = ',', encoding = 'UTF-8')
resale_data_20122014 <- read.csv('resale_Mar2012toDec2014.csv', header = TRUE, sep = ',', encoding = 'UTF-8')
resale_data_2017 <- read.csv('resale_Jan2017onwards.csv', header = TRUE, sep = ',', encoding = 'UTF-8')

resale_data_19901999
resale_data_20002012
resale_data_20122014
resale_data_2017
```

Next, we notice that 3 of 4 dataframes have the same columns, we will try to merge these 3 data frames by stacking them on top vertically.

```{r}
resale_data_19902014 <- rbind(resale_data_19901999, resale_data_20002012, resale_data_20122014)

resale_data_19902014
summary(resale_data_19902014)
```

The last data frame has an extra column, which is the "**remaining_lease"** column. We can calculate this value for the rest of the data set, by calculating the number of years between the lease expiry date (by taking the lease_commencement date + 99, since all HDB flats have a 99 year lease) and the date of transaction.

```{r}
resale_data_19902014$remaining_lease <- resale_data_19902014$lease_commence_date + 99 - as.integer(substr(resale_data_19902014$month, start=1, stop=4))

resale_data_19902014
```

The "**remaining_lease"** column in the last data set has type \<chr\>, we will convert it to \<int\> for uniformity before merging.

```{r}
resale_data_2017$remaining_lease <- as.integer(substr(resale_data_2017$remaining_lease, start=1, stop=2))

resale_data_2017
```

We can then merge all 4 datasets into one uniform dataset.

```{r}
resale_data <- rbind(resale_data_19902014, resale_data_2017)
summary(resale_data)

arrange(resale_data, remaining_lease, desc=F)
```

Next, we can remove some of the columns which are irrelevant for our study.

```{r}
clean_resale_data <- resale_data %>%
  select(-block, -street_name, -flat_model)

clean_resale_data
```

```{r}
summary(clean_resale_data)
```

To facilitate more efficient memory usage, stronger data integrity and modelling exploration later on, we will convert columns with \<chr\> data types into factors, since they all represent categorical data.

For **"town"**:

```{r}
clean_resale_data$town <- as.factor(clean_resale_data$town)

count(clean_resale_data, clean_resale_data$town)
```

For **"flat_type"**:

```{r}
count(clean_resale_data, clean_resale_data$flat_type)

# Replace all occurrences of "MULTI GENERATION" in the "flat_type" column with "MULTI-GENERATION"
clean_resale_data$flat_type <- sub("MULTI GENERATION", "MULTI-GENERATION", clean_resale_data$flat_type)

clean_resale_data$flat_type <- as.factor(clean_resale_data$flat_type)
count(clean_resale_data, clean_resale_data$flat_type)
```

For **"storey_range"**:

```{r}
count(clean_resale_data, clean_resale_data$storey_range)

# Not sure how to handle the overlaps LMAO @marcus do you think we can fix these into categorical variables?
```

```{r}
summary(clean_resale_data)
```

## Exploratory Data Analysis for Resale Flat Prices

### Overall Time Series Analysis of Prices (marcus)

#testing git upload - marcus

### Distribution of Price by Flat Type (marcus)

### Distribution of Price by Town (zz)

-   Heat-map representation, over-time?

### Distribution of Price by Remaining Lease Years (marcus)

-   Predict the future price of HDB for Remaining lease years of less than 40 (or whatever the lowest remaining lease years is now)

### Distribution of Price by Storey range of flat (marcus)

\

## Modelling and Prediction of future HDB Resale Prices (xf)

-   Linear regression for single variables

-   Linear regression for 2 variables

-   Linear regression for all variables

\

## How BTO Buyers can maximise returns (KIV)

-   Calculate returns based on:

    -   Number of years after minimum MOP

    -   Town

    -   Room Type
