---
title: "finalgroupproject"
author: "Tee Zhi Zhang"
date: "2023-11-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Final Group Project

### Group: Marcus Wan, Tan Xue Feng, Tee Zhi Zhang

## Introduction of Data

-   Price of HDB Resale Flat Sales: <https://beta.data.gov.sg/collections/189/datasets/d_8b84c4ee58e3cfc0ece0d773c8ca6abc/view>

-   Price of HDB BTO Launches:

    <https://beta.data.gov.sg/datasets/d_2d493bdcc1d9a44828b6e71cb095b88d/view> 

```{r}

library(tidyverse)
library(dplyr)
library(lubridate)
library(sf)
library(gganimate)
library(animation)
library(gifski)
```

## Data Cleaning and Preprocessing for Resale Flat Prices

-   Remove street name, lease commence date, flat model (marcus)

Our data set is split into 5 different csv files which contain data for different time periods. Let us read the data examine the columns and data types.

```{r}
resale_data_19901999 <- read.csv('resale_19901999.csv', header = TRUE, sep = ',', encoding = 'UTF-8')
resale_data_20002012 <- read.csv('resale_2000Feb2012.csv', header = TRUE, sep = ',', encoding = 'UTF-8')
resale_data_20122014 <- read.csv('resale_Mar2012toDec2014.csv', header = TRUE, sep = ',', encoding = 'UTF-8')
resale_data_20152016 <- read.csv('resale_20152016.csv', header=TRUE, sep = ',', encoding = 'UTF-8')
resale_data_2017 <- read.csv('resale_Jan2017onwards.csv', header = TRUE, sep = ',', encoding = 'UTF-8')

resale_data_19901999
resale_data_20002012
resale_data_20122014
resale_data_20152016
resale_data_2017
```

Next, we notice that 3 of 5 dataframes have the same columns, we will try to merge these 3 data frames by stacking them on top vertically.

```{r}
resale_data_19902014 <- rbind(resale_data_19901999, resale_data_20002012, resale_data_20122014)

resale_data_19902014
summary(resale_data_19902014)
```

The last data frame has an extra column, which is the "**remaining_lease"** column. We can calculate this value for the rest of the data set, by calculating the number of years between the lease expiry date (by taking the lease_commencement date + 99, since all HDB flats have a 99 year lease) and the date of transaction.

```{r}
resale_data_19902014$remaining_lease <- resale_data_19902014$lease_commence_date + 99 - as.integer(substr(resale_data_19902014$month, start=1, stop=4))

resale_data_19902014
```

The "**remaining_lease"** column in the last 2 data set has type \<chr\>, we will convert it to \<int\> for uniformity before merging.

```{r}
resale_data_20152016$remaining_lease <- as.integer(substr(resale_data_20152016$remaining_lease, start=1, stop=2))

resale_data_2017$remaining_lease <- as.integer(substr(resale_data_2017$remaining_lease, start=1, stop=2))

resale_data_20152016
resale_data_2017
```

We can then merge all 5 datasets into one uniform dataset.

```{r}
resale_data <- rbind(resale_data_19902014, resale_data_20152016, resale_data_2017)
summary(resale_data)

arrange(resale_data, remaining_lease, desc=F)
```

Next, we can remove some of the columns which are irrelevant for our study.

```{r}
clean_resale_data <- resale_data %>%
  select(-flat_model)

clean_resale_data
```

```{r}
summary(clean_resale_data)
```

To facilitate more efficient memory usage, stronger data integrity and modelling exploration later on, we will convert columns with \<chr\> data types into factors, since they all represent categorical data.

For **"town"**:

```{r}
clean_resale_data$town <- as.factor(clean_resale_data$town)

count(clean_resale_data, clean_resale_data$town)
```

For **"flat_type"**:

```{r}
count(clean_resale_data, clean_resale_data$flat_type)

# Replace all occurrences of "MULTI GENERATION" in the "flat_type" column with "MULTI-GENERATION"
clean_resale_data$flat_type <- sub("MULTI GENERATION", "MULTI-GENERATION", clean_resale_data$flat_type)

clean_resale_data$flat_type <- as.factor(clean_resale_data$flat_type)
count(clean_resale_data, clean_resale_data$flat_type)
```

For **"storey_range"**:

```{r}
count(clean_resale_data, clean_resale_data$storey_range)

# Since the storey data is provided in a range, we shall determine the average storey and use that for analysis.
clean_resale_data <- clean_resale_data %>%
  mutate(storey_min = as.numeric(sub(" TO .*", "", storey_range)),
         storey_max = as.numeric(sub(".* TO ", "", storey_range)),
         avg_storey = (storey_min + storey_max) / 2)

```

Changing the date of sale to the correct date format,

```{r}
clean_resale_data$month <- as.Date(paste(clean_resale_data$month, "01", sep = "-"), format = "%Y-%m-%d")
```

```{r}
summary(clean_resale_data)
```

## Exploratory Data Analysis for Resale Flat Prices

### Overall Time Series Analysis of Prices (marcus)

In order to have a better understanding of how the prices of HDB units in general changed over time, we shall conduct a time series analysis of prices.

```{r}
grouped_clean_resale_data <- clean_resale_data %>%
  group_by(month) %>%
  summarize(avg_price = mean(resale_price, na.rm = TRUE))

ggplot(grouped_clean_resale_data, aes(x = month, y = avg_price)) +
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Resale Price Over Time", x = "Time", y = "Resale Price")
```

### Distribution of Price by Flat Type (marcus)

We can also analyse the data to analyse the prices of the different types of HDB flats.

```{r}
summary_stats <- clean_resale_data %>%
  group_by(flat_type) %>%
  summarize(
    Mean_Price = mean(resale_price),
    Median_Price = median(resale_price),
    Min_Price = min(resale_price),
    Max_Price = max(resale_price)
  )
summary_stats
```

Visualising using Box Plot,

```{r}
ggplot(clean_resale_data, aes(x = flat_type, y = resale_price)) +
  geom_boxplot() +
  labs(title = "Distribution of Prices by Flat Type", y = "Resale Price (SGD)", x = "flat type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Visualising using Bar Plot,

```{r}
ggplot(summary_stats, aes(x = flat_type, y = Mean_Price)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(title = "Average Price by Flat Type", y = "Average Resale Price (SGD)", x = "flat type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::comma)

```

Visualizing using Histogram,

```{r}
ggplot(clean_resale_data, aes(x = resale_price)) +
  geom_histogram(binwidth = 50000, fill = "green", color = "black") +
  facet_wrap(~flat_type, scales = "free") +
  labs(title = "Price Distribution by Flat Type", x = "Resale Price (SGD)", x = "flat type") +
  scale_x_continuous(labels = scales::comma)

```

As noticed in the time-series analysis, the prices of HDB flats changed greatly over time. It might thus be more useful to compare the price of various flat types over a shorter time frame. We chose 2017 for this as this is the latest data.

```{r}
clean_resale_data_2017 <- clean_resale_data %>%
  filter(year(month) == 2017)
```

```{r}
summary_stats_2017 <- clean_resale_data_2017 %>%
  group_by(flat_type) %>%
  summarize(
    Mean_Price = mean(resale_price),
    Median_Price = median(resale_price),
    Min_Price = min(resale_price),
    Max_Price = max(resale_price)
  )
summary_stats_2017
```

Visualising using Box Plot,

```{r}
ggplot(clean_resale_data_2017, aes(x = flat_type, y = resale_price)) +
  geom_boxplot() +
  labs(title = "Distribution of Prices by Flat Type", y = "Resale Price (SGD)", x = "flat type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Visualising using Bar Plot,

```{r}
ggplot(summary_stats_2017, aes(x = flat_type, y = Mean_Price)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(title = "Average Price by Flat Type", y = "Average Resale Price (SGD)", x = "flat type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::comma)

```

Visualizing using Histogram,

```{r}
ggplot(clean_resale_data_2017, aes(x = resale_price)) +
  geom_histogram(binwidth = 50000, fill = "green", color = "black") +
  facet_wrap(~flat_type, scales = "free") +
  labs(title = "Price Distribution by Flat Type", x = "Resale Price (SGD)", x = "flat type") +
  scale_x_continuous(labels = scales::comma)

```

### Distribution of Price by Town (zz)

In order to get a mapped representation of all our data points, we needed to convert the addresses to latitude, longitude coordinates. We can do so by using geocoding services such as Google Maps', but this will be very costly as we have close to 900k address to encode. Thankfully, we found a mapping dataset on Kaggle: <https://www.kaggle.com/datasets/mylee2009/singapore-postal-code-mapper/> that allowed us to do this mapping.

Here, we read and renamed the column names to match the column names of our data frame and selected only the relevant columns, to prepare for inner joining.

```{r}
locations_df <- read_csv("sg_zipcode_mapper.csv")
locations_df
locations_df <- locations_df %>%
  rename("street_name" = "RD_name", "block" = "blk_no") %>%
  select(street_name, latitude, longtitude, block)
```

Next, we did an inner join to merge information about the coordinates on to our original dataframe.

Our initial results only showed 26k matched rows out of over 900k rows, which was puzzling for us. Hence we examined the location-mapping dataframe closer with *View(locations_df)* and found out that there were different naming conventions for the street names. Hence, we rectified that manually in the .csv files, and our matches increased to over 700k rows, which was sufficient.

```{r}
clean_resale_data
```

```{r}

matched_rows <- inner_join(clean_resale_data, locations_df, by = c("street_name", "block"))
matched_rows$price_psm <- matched_rows$resale_price / matched_rows$floor_area_sqm

count(matched_rows, matched_rows$town)
matched_rows
```

```{r}
# We found and used a vector file that makes up Singapore's map.
singapore_map <- st_read("singapore-boundary.geojson")

# Adding labels of each town to the map:
labels <- data.frame(
  longitude = c(103.847641, 103.932381, 103.848747, 103.818588, 103.7630947, 103.7763724, 103.84622192382812, 103.74625396728516, 103.76336669921875, 103.8818588256836, 103.89112091064453, 103.74030303955078, 103.70392608642578, 103.85871887207031, 103.90089416503906, 103.95957946777344, 103.90625762939453, 103.7967529296875, 103.82157135009766, 103.89371490478516),
  latitude = c(1.371730 ,1.324660, 1.350400, 1.282010, 1.3779175, 1.3546901, 1.2873523235321045, 1.3849343061447144, 1.3065141439437866, 1.3184434175491333, 1.3708921670913696, 1.3372708559036255, 1.3410615921020508, 1.3161046504974365, 1.3020074367523193, 1.3678756952285767, 1.4028823375701904, 1.3026505708694458, 1.4498075246810913, 1.3933453559875488), 
  label = c("Ang Mo Kio", "Bedok", "Bishan", "Bukit Merah", "Bukit Panjang", "Bukit Timah", "Central", "Choa Chu Kang", "Clementi", "Geylang", "Hougang", "Jurong East", "Jurong West", "Kallang", "Marine Parade", "Pasir Ris", "Punggol", "Queenstown", "Sembawang", "Sengkang"
))

# Next we did a ggplot with the Singapore map as the underlying background layer, with points positioned based on their lat-long coordinates, and colored by their price per square metre.

base_plot <- ggplot() +
  geom_sf(data = singapore_map) +
  coord_sf(crs = 4326, xlim = c(103.67, 104.0), ylim = c(1.26, 1.465), expand = FALSE) +
  geom_point(data = matched_rows, aes(x = longtitude, y = latitude,  color= price_psm), size = 0.5) +
  scale_color_gradient(low = "green", high = "red") +
  geom_text(data = labels, aes(x = longitude, y = latitude, label = label), size = 3) +
  labs(title = "Singapore Map with Data Points") +
  theme_minimal()

base_plot
```

```{r}
animation <- base_plot +
  transition_time(month) +
  labs(title = "Month: {frame_time}")

animate(animation, duration = 10, fps = 6, width = 600, height = 300, renderer = gifski_renderer())
```

Here, we can see that property prices have been increasing steadily since the 1990s, starting from towns in the South like Bukit Merah and Central, and then the increase gradually expanded to other towns like Ang Mo Kio, Punggol and Bishan.

Next, let us explore the number of transactions across towns:

```{r}
monthly_count <- clean_resale_data %>%
  group_by(month) %>%
  summarize(Count = n()) %>%
  arrange(Count, DESC=T)

monthly_count

ggplot(monthly_count, aes(x = month, y = Count)) +
  geom_line() +
  labs(
    title = "Number of Transactions Over Time",
    x = "Month",
    y = "Transaction Count"
  )
```

From this, we observe that there is a large spike in the number of transactions in the years 1998-2000, where Singapore experienced a property "boom" period; and we also can see that the build-up and gradual stabilization took around 3-4 years each. There are smaller peaks in the years of 2010 and 2021 as well. Overall, we can see that the trend follows a wave pattern where there are repeated periods of ups and downs that span around 10-15 years each.

```{r fig.width=10, fig.height=6}

ggplot(data = clean_resale_data) +
  geom_bar(aes(x = month, color=town))
  facet_wrap(clean_resale_data$town) +
  labs(title = "Number of Transactions Over Time",
       x = "Year-Month",
       y = "Transaction Count") +
  theme(legend.key.size = unit(1, "lines"))
```

```{r}
monthly_count <- clean_resale_data %>%
  group_by(month, town) %>%
  summarize(Count = n()) %>%
  arrange(Count, DESC=T)

monthly_count

ggplot(monthly_count, aes(x = month, y = Count, color=town)) +
  geom_line() +
  labs(
    title = "Number of Transactions Over Time",
    x = "Month",
    y = "Transaction Count"
  )
```

We can observe that the peak in 1998-2000 comprised of all-time high number of HDB resales happening in Bedok, Hougang, Pasir Ris, Jurong West, Tampines, Yishun; while the peak in 2010 can be attributed to the towns of Jurong West, Woodlands, Yishun, Bedok and Clementi. The most recent peak in 2022 can be mainly attributed to Yishun, Woodlands, Choa Chu Kang, Tampines and Jurong West.

Next, let's explore how the resale prices vary across HDB Towns.

We first see the overall resale prices trend in Singapore:

```{r}
ggplot(data = clean_resale_data, aes(x=month, y=resale_price)) +
  geom_point() +
  geom_smooth() +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Resale Price Over Time", x = "Time", y = "Resale Price")

```

Next, we look at the trend in each HDB Town:

```{r fig.width=10, fig.height=6}
ggplot() +
  geom_smooth(data = clean_resale_data, aes(x = month, y = resale_price, color=town)) +
  facet_wrap(clean_resale_data$town) +
  scale_y_continuous(labels = scales::comma) +
  theme(axis.text.x = element_blank()) +
  labs(title = "Resale Price Over Time for Various Towns", x = "Time", y = "Resale Price")
```

We can see that each town generally follows the overall trend of Singapore's HDB resale prices: with a steady increase over time but with periodic ups and downs in property prices. It can be noted that the peak and trendline for property prices typically trail those for the number of transactions by 1-2 years, shedding light on Singapore's property market behavior.

There is a lack of data for Lim Chu Kang town because in Lim Chu Kang, there was only a small public housing area known as Neo Tiew Estate (or Lim Chu Kang Rural Centre). The whole area was en-bloc in 2002, with the residents shifted to Jurong West. The estate was vacated since then, and is currently used for FIBUA (Fighting in Built-Up Areas) trainings by the Singapore Armed Forces (SAF). Hence, there is little to no transactions happening in Lim Chu Kang.

##punggol

```{r}
summary_town <- clean_resale_data %>%
  group_by(town) %>%
  summarise(Mean_Price = mean(resale_price))

ggplot(data = summary_town, aes(x = town, y= Mean_Price)) +
  geom_bar(stat = "identity", fill="blue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Resale Price for Various Towns", x = "Town", y = "Mean Resale Price")
```

```{r fig.width=10, fig.height=8}
# Consider only the most recent data, after 2020: 
filtered_data <- clean_resale_data %>%
  filter(year(month) >= 2020)

ggplot() +
  geom_boxplot(data = filtered_data, aes(y = resale_price)) +
  facet_wrap(filtered_data$town)
```

Here, we can get a bird's eye view of the relative pricing of each town's HDB properties through the median resale price, 1st and 3rd quartile price and the minimum/ maximum prices.

You would have to pay a premium for flats in towns like Bishan, Bukit Merah, Bukit Timah, Central Area, Queenstown and Toa Payoh with median prices of around \$750K; primarily due to their locations and proximity to workplaces and places of attraction. On the other end, flats in towns like Choa Chu Kang, Bukit Panjang, Sembawang, Yishun and Pasir Ris are more affordable with median prices of around \$400K.

### Distribution of Price by Remaining Lease Years (marcus)

As HDB units is a form of public housing, all HDB units sold to Singaporeans have a 99 year lease attached to them. Once the period of the lease is up, the unit must be returned to the government without compensation (ie, its value will be 0 when the lease hits 0 years).

Thus, one factor that affects the price of a HDB unit is the number of years of lease left. We shall analyse this here.

```{r}
remaining_lease_summary_stats <- clean_resale_data %>%
  group_by(remaining_lease) %>%
  summarize(
    Mean_Price = mean(resale_price),
    Median_Price = median(resale_price),
    Min_Price = min(resale_price),
    Max_Price = max(resale_price)
  )
remaining_lease_summary_stats
```

Scatter plot

```{r}
#grouped_remaining_lease <- clean_resale_data %>%
#  group_by(remaining_lease) %>%
#  summarize(avg_price = mean(resale_price, na.rm = TRUE))

#ggplot(grouped_remaining_lease, aes(x = remaining_lease, y = avg_price)) +
#  geom_line() +
#  scale_y_continuous(labels = scales::comma) +
#  labs(title = "Average Price vs Remaining Lease", x = "Remaining Lease", y = "Resale Price")

ggplot(clean_resale_data, aes(x = remaining_lease, y = resale_price)) +
  geom_point() +
  geom_smooth() +
  labs(title = "Scatterplot of Resale Prices vs. Remaining Lease", x = "Remaining Lease (Years)", y = "Resale Price (SGD)")
```

#This data is too dirty. In order to ensure that we see a clearer relationship between the price and the remaining lease, let's only use data from 2017.

```{r}
remaining_lease_summary_stats_2017 <- clean_resale_data_2017 %>%
  group_by(remaining_lease) %>%
  summarize(
    Mean_Price = mean(resale_price),
    Median_Price = median(resale_price),
    Min_Price = min(resale_price),
    Max_Price = max(resale_price)
  )
remaining_lease_summary_stats_2017
```

Scatter plot

```{r}
ggplot(clean_resale_data_2017, aes(x = remaining_lease, y = resale_price)) +
  geom_point() +
  geom_smooth() +
  labs(title = "Scatterplot of Resale Prices vs. Remaining Lease", x = "Remaining Lease (Years)", y = "Resale Price (SGD)")
```

Using a regression model,

```{r}
lm_model_2017 <- lm(resale_price ~ remaining_lease, data = clean_resale_data_2017)

# Summary of the regression model
summary(lm_model_2017)
```

Visualizing the regression line, #remove

```{r}
ggplot(clean_resale_data_2017, aes(x = remaining_lease, y = resale_price)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Regression Line: Resale Prices vs. Remaining Lease", x = "Remaining Lease (Years)", y = "Resale Price (SGD)")
```

We can see that there is a positive relationship between remaining lease and the resale price.

### Predict the future price of HDB for Remaining lease years of less than 40 (or whatever the lowest remaining lease years is now)

From the sales data we are utilising, we can see that the lowest remaining lease value is 42 years. This is because it has only been around 60 years since the first HDB unit was sold. As such, til date, there has been no unit that was returned to the government. We know that the price of these units will be zero when the lease runs out. As such, one point of interest we can analyse would be whether this fact is taken into account by the resale market.

Let's create an additional data point for when resale price = 0 when remaining lease = 0. We shall create a new weighted regression model with increased weight for the (0, 0) data point.

```{r}
additional_point <- data.frame(
  remaining_lease = 0,
  resale_price = 0
)
clean_resale_data_2017_00 <- bind_rows(clean_resale_data_2017, additional_point)

weights <- ifelse(clean_resale_data_2017_00$remaining_lease == 0, 20509, 1)

weighted_lm_model <- lm(resale_price ~ remaining_lease, data = clean_resale_data_2017_00, weights = weights)

summary(weighted_lm_model)

```

Visualizing on scatter plot,

```{r}
# Scatterplot of data
plot(clean_resale_data_2017_00$remaining_lease, clean_resale_data_2017_00$resale_price, pch = 19, col = "black", main = "Weighted Regression Line: Resale Prices vs. Remaining Lease", xlab = "Remaining Lease (Years)", ylab = "Resale Price (SGD)")

# Add the fitted line from the regression model
abline(weighted_lm_model, col = "red")
```

Price Prediction Data Using Locally estimated scatterplot smoothing,

```{r}
ggplot(clean_resale_data_2017_00, aes(x = remaining_lease, y = resale_price)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "Scatterplot of Resale Prices vs. Remaining Lease", x = "Remaining Lease (Years)", y = "Resale Price (SGD)")
```

Price prediction using ggtrendline ##not working, cant combine

```{r}
library(ggtrendline)
ggtrendline(clean_resale_data_2017_00$remaining_lease, clean_resale_data_2017_00$resale_price, model = "line3P")
ggplot(clean_resale_data_2017_00, aes(x = remaining_lease, y = resale_price)) +
  geom_point() +
  labs(title = "Scatterplot of Resale Prices vs. Remaining Lease", x = "Remaining Lease (Years)", y = "Resale Price (SGD)")
```

Prediction of price trend using forecast, ##not working

```{r}
library(forecast)

aggregated_data <- clean_resale_data_2017%>%
  group_by(remaining_lease) %>%
  summarize(avg_resale_price = mean(resale_price))

# Convert data to time series
ts_data <- ts(aggregated_data$avg_resale_price)

# Fit an ARIMA model (replace with your model)
arima_model <- auto.arima(ts_data)

# Make forecasts for years 1 to 41
forecast_values <- forecast(arima_model, h = 41)

# Create a data frame for visualization
df <- data.frame(
  year = 48:96,
  actual = c(0, aggregated_data$avg_resale_price),
  forecast = forecast_values$mean
)

# Create a ggplot to visualize historical and forecasted data
ggplot(df, aes(x = year, y = actual)) +
  geom_line(aes(color = "Actual")) +
  geom_line(aes(x = year, y = forecast, color = "Forecast")) +
  labs(title = "Historical and Forecasted Price Trend", x = "Year", y = "Resale Price") +
  scale_color_manual(values = c("Actual" = "blue", "Forecast" = "red"))

```

Predict trend using least square, ##ugly af, dk how fix

```{r}

additional_point <- data.frame(
  remaining_lease = 0,
  resale_price = 0
)
clean_resale_data_2017_00 <- bind_rows(clean_resale_data_2017, additional_point)

# Define a nonlinear model (you need to choose an appropriate nonlinear function)
polynomial_model <- lm(resale_price ~ remaining_lease + I(remaining_lease^2), 
                      data = clean_resale_data_2017_00)

# Make predictions with the nonlinear model
predictions <- predict(polynomial_model, newdata = data.frame(remaining_lease = 1:99))

# Create a data frame for visualization
df <- data.frame(
  remaining_lease = c(clean_resale_data_2017_00$remaining_lease, 1:99),
  resale_price = c(clean_resale_data_2017_00$resale_price, predictions)
)

# Create a ggplot to visualize the data and nonlinear best-fit line
ggplot(df, aes(x = remaining_lease, y = resale_price)) +
  geom_point(data = clean_resale_data_2017_00, color = "blue") +
  geom_smooth(data = df, aes(x = remaining_lease, y = resale_price), color = "red") +
  labs(title = "Nonlinear Best-Fit Line", x = "Remaining Lease", y = "Resale Price")
```

### Distribution of Price by Storey range of flat (marcus)

Let's also explore how the storey of the unit affects the price

```{r}
# Since there are too many storeys to display on a barplot, let's create bins
bin_width <- 10 
clean_resale_data_bins <- clean_resale_data %>%
  mutate(storey_bin = cut(avg_storey, breaks = seq(0, max(avg_storey), by = bin_width), right = FALSE))

storey_bin_summary_stats <- clean_resale_data_bins %>%
  group_by(storey_bin) %>%
  summarize(
    Mean_Price = mean(resale_price),
    Median_Price = median(resale_price),
    Min_Price = min(resale_price),
    Max_Price = max(resale_price)
  )
storey_bin_summary_stats 
```

Visualising using Box Plot, ##remove NA

```{r}
ggplot(clean_resale_data_bins, aes(x = storey_bin, y = resale_price)) +
  geom_boxplot() +
  labs(title = "Distribution of Prices by Storey Range", y = "Resale Price (SGD)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Visualising using a barplot,

```{r}
ggplot(storey_bin_summary_stats, aes(x = storey_bin, y = Mean_Price)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(title = "Average Price by Storey Range", x = "Storey Range", y = "Average Resale Price (SGD)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Visualising using a histogram,

```{r}
ggplot(clean_resale_data_bins, aes(x = resale_price)) +
  geom_histogram(binwidth = 50000, fill = "green", color = "black") +
  facet_wrap(~storey_bin, scales = "free") +
  labs(title = "Price Distribution by Storey Range", x = "Resale Price (SGD)")

```

## Modelling and Prediction of future HDB Resale Prices (xf)

-   Linear regression for 2 variables

-   Linear regression for all variables

\

## How BTO Buyers can maximise returns (KIV)

HDB BTO (Build-To-Order) flats are a type of public housing in Singapore, offered by the Housing and Development Board (HDB). These flats are designed to cater to the housing needs of Singaporean citizens and are typically more affordable than private housing options.

The Minimum Occupancy Period (MOP) is the minimum duration that homeowners are required to occupy their BTO flats before they can sell or rent them. The MOP duration typically ranges from 5 to 10 years, depending on the flat type and whether the flat was bought under the singles scheme. For example, the MOP for a standard family nucleus in a 4-room flat is typically 5 years.

```{r}
BTO_data <- read_csv("BTO.csv")

BTO_town_counts <- count(BTO_data ,BTO_data$town)
BTO_data
BTO_town_counts
```

We will select only projects with more than 20 BTO launches in this dataset, to make sure our data and analysis are significant. We will also remove rows where the selling price is 0, meaning to say that there was no launch of that room-type in the town.

```{r}
BTO_data_merged <- merge(BTO_data ,BTO_town_counts, by.x = "town", by.y = "BTO_data$town")
  
BTO_data_filtered <- filter(BTO_data_merged ,BTO_data_merged$n > 20)

BTO_data_final <- filter(BTO_data_filtered, BTO_data_filtered$min_selling_price > 0)
```

Next, we further clean this data to make sure that the columns of town and room-types and their corresponding values match that of the clean_resale_data figures.

```{r}
BTO_data_final$town_upper = toupper(BTO_data_final$town)

mapping <- c("2-room" = "2 ROOM", "3-room" = "3 ROOM", "4-room" = "4 ROOM", "5-room" = "5 ROOM")

BTO_data_final <- BTO_data_final %>%
  mutate(flat_type = case_when(
    room_type %in% names(mapping) ~ mapping[room_type],
    TRUE ~ NA_character_
  ))

BTO_data_final
count(BTO_data_final, BTO_data_final$flat_type)

```

Since BTOs have a minimum occupancy period of 5 years, and the earliest BTO project we have in the data set is in 2008, we will only require resale prices from year 2013 onwards, and we do the filtering. Next, we got the mean resale price for each town, flat type and year from clean_resale_data, to be merged later on with BTO_data_final.

```{r}
resale_data_for_bto <- clean_resale_data %>%
  mutate(year = year(month)) %>%
  group_by(year, town, flat_type) %>%
  filter(town %in% c("CHOA CHU KANG", "PUNGGOL", "SEMBAWANG", "SENGKANG", "WOODLANDS", "YISHUN")) %>%
  filter(year >= 2013) %>%
  summarise(mean_resale_price = mean(resale_price))

resale_data_for_bto

```

We want to see what the mean resale price of each BTO listing, every year after the MOP is up (assuming that the MOP is 5 years for all of these BTO projects). We look 10 years after the MOP is up, and add 10 columns for every row to the BTO_final_data. We then populate these columns by checking for the year, town and room-type data and getting the mean_resale_price value from resale_data_for_bto:

```{r}
for (i in 1:10) {
  column_name <- paste("Mean Resale Price ", i, " years after the MOP")
  
  mean_resale_price_values <- numeric(nrow(BTO_data_final))
  
  for (j in 1:nrow(BTO_data_final)) {
    year_condition <- BTO_data_final$financial_year[j] + i + 5 == (resale_data_for_bto$year) &
                     BTO_data_final$town_upper[j] == resale_data_for_bto$town &
                    BTO_data_final$flat_type[j] == resale_data_for_bto$flat_type
    if (any(year_condition)) {
      mean_resale_price_values[j] <- (resale_data_for_bto$mean_resale_price[year_condition])
    } else {
      mean_resale_price_values[j] <- NA
    }
  }
  
  BTO_data_final[[column_name]] <- mean_resale_price_values
}

view(BTO_data_final)

BTO_data_final
```

Calculate returns based on:

### Number of years after minimum MOP

In this section, we want to find out if there is an optimal time for BTO owners to sell their flats to the public market in order to maximize profits from the sale.

We only have full 10 years of data from BTO projects launched in 2008: Punggol, Sengkang and Woodlands. We will also standardise the flat-type to 4-rooms, and use those as our data for this study.

```{r}
rows_years <- subset(BTO_data_final, financial_year == 2008 & flat_type == "4 ROOM")

rows_years
```

We will then calculate the potential profit in each year after the MOP is up. To standardize the purchase price, we will take the mean between the min_selling_price and max_selling_price.

```{r}
rows_years %>%
  mutate(mean_purchase_price = (min_selling_price+max_selling_price) / 2)

profit_over_years <- rows_years %>%
  rename_with(~ str_replace_all(., "Price", "Profit"), everything())


profit_over_years
rows_years
```

-   Town

-   Room Type
